Optimal Peak Shaving with Bifacial Solar PV: Simulated Storage Battery Dispatch for Electric Vehicle Charging Load

### 2024-01-09 Paper/paper.tex
few paper edits

Even in cases where the electric load isn't growing, solar photovoltaic (PV) arrays that are large relative to the load will create relative peaks and valleys in the load seen by the network as clouds pass overhead. As the solar ramp-down rate due to clouds can be much faster than the ramp-up rate of many traditional thermal generators, grid operators will need to predict and anticipate these relative peaks. With an all-sky camera and machine vision algorithms, 15-minute nowcasting cloud identification may only reach 84.2\% accuracy \cite{Nespoli2022}. The remaining peaks will likely need to be managed by other means.

@@

Peak load management, or peak shaving, is the process of reducing the maximum load at a given point in an electrical network. This will necessarily increase the load factor ($P_{load,avg} P_{load,max}^{-1}$) even if load curtailment is the only method used. Other methods include load scheduling, load shifting using energy storage (battery or otherwise), and reduction of the effective load with distributed generation (DG) \cite{Uddin2018}. Peak shaving may include the concept of a peak load threshold which defines a maximum desired peak load during a period, typically on the order of hours. Grid operators may use peak shaving to increase power quality, defer future infrastructure upgrades, decrease losses, or reduce energy costs. End-users typically reduce their peaks solely on a cost basis. Lastly, some carbon reduction is possible by reducing the on/off transitions of peaker plants due to load peaks \cite{Uddin2018}.

Broadly speaking there are two categories of peak shaving with respect to the goal: technical and economic. Some major differences between the two are summarized in Table \ref{tab:econ-tech-peak-shaving}.

@@

Economic peak shaving aims to reduce what a consumer pays for power and or energy. Medium and large electric consumers often pay an energy price ($p_e\ [\$/kWh]$) and a power price ($p_p\ [\$/kW]$), also called a demand charge. Each may vary with time of day, day of week, and season of the year, which is often referred to as time of use (TOU) or peak pricing. The power cost ($p_p P_{load,max}\ [\$]$) typically applies to the max power during the billing period, where the max power is the maximum non-moving average calculated on a given interval, such as 60 minutes. Where there is a sufficient spread between the peak and off-peak prices there may be the opportunity to reschedule, shift, or curtail load, or dispatch DG.

In the literature, peak shaving papers often don't consider the retail cost of energy and rarely utilize vertical bifacial solar PV or a threshold-based approach. In \cite{Reihani2016} authors forecast load and choose a state of charge (SOC) trajectory to manage load peaks. Linear programming optimization in \cite{Lee2014} reduces the on-peak load of residential EV charging by dispatching a battery. In \cite{Gazafroudi2018} a multi-agent system buys and sells energy in a local energy market for a residence with solar PV and a dispatchable vehicle-to-home EV battery. In \cite{Zheng2015} authors optimize peak shaving based on a rate tariff and a technical power threshold. Vertical bifacial has been benchmarked in the literature and evaluated for new applications such as public highway solar installations \cite{Niccolai2022}. Vertical bifacial solar is coupled with a storage battery in \cite{Wallberg2022} and in \cite{Castellucci2022} with a rule-based peak shaving algorithm, reducing EV charging load peaks by 38.6\% and 79\%, respectively. However, the retail energy cost isn't estimated and a comparison is not made to understand the relative merit of vertical bifacial compared to south-facing modules. Bifacial PV has been studied extensively in the literature, performing around 11\% better than comparable monofacial PV \cite{Ogliari2023}, .

@@

A stationary, storage, AC battery system is simulated as the only dispatchable asset in the peak shaving algorithm. In each time step the battery is chosen to charge or discharge within its technical limits. Because the control action is to hold the grid power exchange below a chosen threshold, batteries with a larger power capacity or energy capacity will necessarily achieve a lower threshold if the entire battery capacity is used. For a given energy capacity (\(kWh\)) those limits are a charge or discharge rate of no more than 1C and a state of charge (SOC) between 0 and 100\%. State of energy (SOE) is the SOC multiplied by nominal energy capacity and is reported in plots because the units are more meaningful when comparing dispatch time series. Half-round-trip charge and discharge efficiencies are constant at 90\%. Self-discharge and thermal limiting are not modeled specifically.

@@

An optimal peak shaving strategy is used to minimize the total retail electric cost to the EV charging station. The strategy is operational only and assumes the solar production time series and battery capacity are fixed. The state variable controlled by the algorithm is battery charge or discharge power. However, the algorithm reframes the problem as one of choosing a power threshold \(T\ [kW]\) applied to the point of injection to the grid, the electric meter. The battery is then dispatched, within its technical limits, to hold the net load ($L-S$) below the threshold. When the algorithm is successful the different thresholds for each TOU period are met for an entire month. If the battery reaches a technical limit and net load exceeds a threshold, the simulation is not necessarily invalid but is not likely to minimize the retail electric cost to the site for that combination of solar and battery size. The peak shaving logic is described in Figure \ref{fig:peakshaving-flowchart}.

\begin{figure}
  \centering
  \includegraphics[width=7.5cm]{./images/peak shaving flowchart.png}
  \caption{Peak Shaving Flowchart. Net load is the time series of load with solar subtracted. There is one threshold value for one for each TOU period with a non-zero power price. For each timestep of the simulation, the net load is compared to the threshold of the current TOU period. The battery is discharged if the net load is greater than the threshold and charged if the net load is less than the threshold. If the two are equal or if there is no power price the battery does nothing.}
  \label{fig:peakshaving-flowchart}
\end{figure}

@@

\begin{table}[!h]
  \caption{Timeseries Data Summary. The primary data sources are Caltech ACN EV charging power and GOES solar irradiance. Solar production is modeled from irradiance.}
  \label{tab:data-summary}
  \begin{tabularx}{\linewidth}{XXXXX}
    \toprule
    Description                     & Location                                & Type                                                                                                                  & Interval                               & Source                                \\
    \midrule
    EV charging power $[kW_{ac}]$   & Jet Propulsion Lab, Pasadena CA, USA    & Measured and aggregated from multiple  EV chargers at single site                                                     & 10 sec (average downsampled to 15 min) & Caltech ACN (site JPL) \cite{Lee2021} \\
    Solar irradiance $[W/m^2]$      & GPS: 34.2013, -118.1721 (2x2 km square) & GOES satellite irradiance                                                                                             & 5 min                                  & NSRDB PSMv3 \cite{Sengupta2018}       \\
    PV array production $[kW_{ac}]$ & GPS: 34.2013, -118.1721 (2x2 km square) & Modelled from GOES satellite irradiance, 368 Prism Solar 350 W bifacial modules, 368 Enphase IQH 380 W microinverters & 15 min                                 & SAM \cite{NREL2022}                   \\
    \bottomrule
  \end{tabularx}
\end{table}

\begin{table}[!h]
  \caption{Retail Electric Tariff. An applicable electric tariff schedule with power and energy prices for several different TOU periods. From California PG\&E.}
  \label{tab:tariff}
  \begin{tabularx}{\textwidth}{XXXXX}
    \toprule
    Name            & TOU Period                         & Summer (Jun 1 - Sept 30) $[\$]$ & Winter (Oct 1 -Feb 28) $[\$]$ & Spring (Mar 1 - May 30) $[\$]$ \\
    \midrule
    All hours       & 0:00-0:00     3                    & 26.07/$kW$                      & 26.07/$kW$                    & 26.07/$kW$                     \\
    Super off-peak  & 9:00-14:00                         & -\/-                            & -\/-                          & 0.079/$kWh$                    \\
    Off-peak spring & 0:00-9:00, 14:00-16:00, 21:00-0:00 & -\/-                            & -\/-                          & 0.132/$kWh$                    \\
    Off-peak winter & 0:00-16:00, 21:00-0:00             & -\/-                            & 0.132/$kWh$                   & -\/-                           \\
    Off-peak summer & 0:00-14:00, 23:00-0:00             & 0.132/$kWh$                     & -\/-                          & -\/-                           \\
    Partial-peak    & 14:00-16:00, 21:00-23:00           & 6.81/$kW$, 0.159/$kWh$          & -\/-                          & -\/-                           \\
    Peak            & 16:00-21:00                        & 32.90/$kW$, 0.196/$kWh$         & 2.22/$kW$, 0.172/$kWh$        & 2.22/$kW$, 0.172/$kWh$         \\
    \bottomrule
  \end{tabularx}
\end{table}

>>>>

Even in cases where the electric load isn't growing, solar Photovoltaic (PV) arrays that are large relative to the load will create relative peaks and valleys in the load seen by the network as clouds pass overhead. As the solar ramp-down rate due to clouds can be much faster than the ramp-up rate of many traditional thermal generators, grid operators will need to predict and anticipate these relative peaks. With an all-sky camera and machine vision algorithms, 15-minute nowcasting cloud identification may only reach 84.2\% accuracy \cite{Nespoli2022}. The remaining peaks will likely need to be managed by other means.

@@

Peak load management, or peak shaving, is the process of reducing the maximum load at a given point in an electrical network. This will necessarily increase the load factor ($P_{load,avg} P_{load,max}^{-1}$) even if load curtailment is the only method used. Other methods include load scheduling, load shifting using energy storage (battery or otherwise), and reduction of the effective load with Distributed Generation (DG) \cite{Uddin2018}. Peak shaving may include the concept of a peak load threshold which defines a maximum desired peak load during a period, typically on the order of hours. Grid operators may use peak shaving to increase power quality, defer future infrastructure upgrades, decrease losses, or reduce energy costs. End-users typically reduce their peaks solely on a cost basis. Lastly, some carbon reduction is possible by reducing the on/off transitions of peaker plants due to load peaks \cite{Uddin2018}.

Broadly speaking there are two categories of peak shaving with respect to the goal, technical and economic. Some major differences between the two are summarized in Table \ref{tab:econ-tech-peak-shaving}.

@@

Economic peak shaving aims to reduce what a consumer pays for power and or energy. Medium and large electric consumers often pay an energy price ($p_e\ [\$/kWh]$) and a power price ($p_p\ [\$/kW]$), also called a demand charge. Each may vary with time of day, day of week, and season of the year, which is often referred to as Time of Use (TOU) or peak pricing. The power cost ($p_p P_{load,max}\ [\$]$) typically applies to the max power during the billing period, where the max power is the maximum non-moving average calculated on a given interval, such as 60 minutes. Where there is a sufficient spread between the peak and off-peak prices there may be the opportunity to reschedule, shift, or curtail load, or dispatch DG.

In the literature, peak shaving papers often don't consider the retail cost of energy and rarely utilize vertical bifacial solar PV or a threshold-based approach. In \cite{Reihani2016} authors forecast load and choose a State of Charge (SOC) trajectory to manage load peaks. Linear programming optimization in \cite{Lee2014} reduces the on-peak load of residential EV charging by dispatching a battery. In \cite{Gazafroudi2018} a multi-agent system buys and sells energy in a local energy market for a residence with solar PV and a dispatchable vehicle-to-home EV battery. In \cite{Zheng2015} authors optimize peak shaving based on a rate tariff and a technical power threshold. Vertical bifacial has been benchmarked in the literature and evaluated for new applications such as public highway solar installations \cite{Niccolai2022}. Vertical bifacial solar is coupled with a storage battery in \cite{Wallberg2022} and in \cite{Castellucci2022} with a rule-based peak shaving algorithm, reducing EV charging load peaks by 38.6\% and 79\%, respectively. However, the retail energy cost isn't estimated and a comparison is not made to understand the relative merit of vertical bifacial compared to south-facing modules. Bifacial PV has been studied extensively in the literature, performing around 11\% better than comparable monofacial PV \cite{Ogliari2023}, .

@@

A stationary, storage, AC battery system is simulated as the only dispatchable asset in the peak shaving algorithm. In each time step the battery is chosen to charge or discharge within its technical limits. Because the control action is to hold the grid power exchange below a chosen threshold, batteries with a larger power capacity or energy capacity will necessarily achieve a lower threshold if the entire battery capacity is used. For a given energy capacity (\(kWh\)) those limits are a charge or discharge rate of no more than 1C and a state of charge (SOC) between 0 and 100\%. State of Energy (SOE) is the SOC multiplied by nominal energy capacity and is reported in plots because the units are more meaningful when comparing dispatch time series. Half-round-trip charge and discharge efficiencies are constant at 90\%. Self-discharge and thermal limiting are not modeled specifically.

@@

An optimal peak shaving strategy is used to minimize the total retail electric cost to the EV charging station. The strategy is operational only and assumes the solar production time series and battery capacity are fixed. The state variable controlled by the algorithm is battery charge or discharge power. However, the algorithm reframes the problem as one of choosing a power threshold \(T\ [kW]\) applied to the point of injection to the grid, the electric meter. The battery is then dispatched, within its technical limits, to hold the net load below the threshold. When the algorithm is successful the different thresholds for each TOU period are met for an entire month. If the battery reaches a technical limit and net load exceeds a threshold, the simulation is not necessarily invalid but is not likely to minimize the retail electric cost to the site for that combination of solar and battery size. The peak shaving logic is described in Figure \ref{fig:peakshaving-flowchart}.

\begin{figure}
  \centering
  \includegraphics[width=7.5cm]{./images/peak shaving flowchart.png}
  \caption{Peak Shaving Flowchart. Net Load (NL) is the time series of load with solar subtracted. There is one threshold value for one for each TOU period with a non-zero power price. For each timestep of the simulation, the net load is compared to the threshold of the current TOU period. The battery is discharged if the net load is greater than the threshold and charged if the net load is less than the threshold. If the two are equal or if there is no power price the battery does nothing.}
  \label{fig:peakshaving-flowchart}
\end{figure}

@@

\begin{table}[!h]
  \caption{Timeseries Data Summary. The primary data sources are Caltech ACN EV charging power and GOES solar irradiance. Solar production is modeled from irradiance.}
  \label{tab:data-summary}
  \begin{tabularx}{\linewidth}{XXXXX}
    \toprule
    Description                     & Location                                & Type                                                                                                                  & Interval                               & Source                                \\
    \midrule
    EV charging power $(kW_{ac})$   & Jet Propulsion Lab, Pasadena CA, USA    & Measured and aggregated from multiple  EV chargers at single site                                                     & 10 sec (average downsampled to 15 min) & Caltech ACN (site JPL) \cite{Lee2021} \\
    Solar irradiance $(W/m^2)$      & GPS: 34.2013, -118.1721 (2x2 km square) & GOES satellite irradiance                                                                                             & 5 min                                  & NSRDB PSMv3 \cite{Sengupta2018}       \\
    PV array production $(kW_{ac})$ & GPS: 34.2013, -118.1721 (2x2 km square) & Modelled from GOES satellite irradiance, 368 Prism Solar 350 W bifacial modules, 368 Enphase IQH 380 W microinverters & 15 min                                 & SAM \cite{NREL2022}                   \\
    \bottomrule
  \end{tabularx}
\end{table}

\begin{table}[!h]
  \caption{Retail Electric Tariff. An applicable electric tariff schedule with power and energy prices for several different TOU periods. From California PG\&E.}
  \label{tab:tariff}
  \begin{tabularx}{\textwidth}{XXXXX}
    \toprule
    Name            & TOU Period                         & Summer (Jun 1 - Sept 30) $(\$)$ & Winter (Oct 1 -Feb 28) $(\$)$ & Spring (Mar 1 - May 30) $(\$)$ \\
    \midrule
    All hours       & 0:00-0:00     3                    & 26.07/$kW$                      & 26.07/$kW$                    & 26.07/$kW$                     \\
    Super off-peak  & 9:00-14:00                         & -\/-                            & -\/-                          & 0.079/$kWh$                    \\
    Off-peak spring & 0:00-9:00, 14:00-16:00, 21:00-0:00 & -\/-                            & -\/-                          & 0.132/$kWh$                    \\
    Off-peak winter & 0:00-16:00, 21:00-0:00             & -\/-                            & 0.132/$kWh$                   & -\/-                           \\
    Off-peak summer & 0:00-14:00, 23:00-0:00             & 0.132/$kWh$                     & -\/-                          & -\/-                           \\
    Partial-peak    & 14:00-16:00, 21:00-23:00           & 6.81/$kW$, 0.159/$kWh$          & -\/-                          & -\/-                           \\
    Peak            & 16:00-21:00                        & 32.90/$kW$, 0.196/$kWh$         & 2.22/$kW$, 0.172/$kWh$        & 2.22/$kW$, 0.172/$kWh$         \\
    \bottomrule
  \end{tabularx}
\end{table}
